Federated Learning's Blessing:\\ FedAvg has Linear Speedup

### 2020-06-01 source/sec_intro.tex

\section{Introduction}
% Motivate FL
The widespread usage of smart home devices and mobile phones constitutes
large-scale distributed networks that generate significant amount of personal data.
Models learned from such data have led to 
% The emergence of those massive amounts of data has empowered machine learning
% solutions to develop 
many successful applications such as personalized
recommendations~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and keyboard predictions~\cite{47586}, etc. Due to the rising awareness of  privacy and the rapid growth of
computational power in mobile devices, it becomes increasingly attractive to
push the computation to the edge. Recent years witnessed the emergence of
federated learning~\cite{mcmahan2016communication} to fulfill this demand.

% Main challenge in FL.
As a distributed learning paradigm, Federated Learning (FL) collaboratively
learns a single statistical model leveraging the data from a massive amount of devices while
keeping the training data isolated.  In particular, FL integrates cross-device
information to achieve improved model quality without transferring training data from individual devices, thus improves user
privacy~\cite{kairouz2019advances}. While promising, however, comparing to the traditional distribution optimization,  FL confronts unique
challenges~\cite{li2019convergence,smith2017federated}, including statistical heterogeneity, that is, the data distributions in different devices can be significantly different, and system heterogeneity, that is, only a subset of devices may be accessible to the central server for model training at each time.

% Gaps in the theoretical understanding. 
To address the aforementioned challenges, Federated
Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as an
effective heuristic which has enjoyed great empirical success~\cite{47586}, attracting
lots of efforts on building up its theoretical understanding 
and accelerated variants. 
Existing convergence guarantees of FedAvg were either largely established on simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or provides a suboptimal 
convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} under the realistic setting is yet to be analyzed.

>>>>

\section{Introduction}
Federated learning (FL) is a machine learning setting where many clients (e.g. mobile devices or whole organizations) collaboratively train a model under the orchestration of a central server (e.g. service provider), while keeping the training data decentralized~\cite{kairouz2019advances}.
In recent years, FL has swiftly emerged to be an important learning paradigm~\cite{mcmahan2016communication}--one that enjoys widespread success in applications such as personalized recommendations~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, keyboard predictions~\cite{47586} to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL. Second, the rising awareness of privacy and the exponential growth of computational power (blessed by Moore's law) in mobile devices have made it increasingly attractive to push the computation to the edge.

Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth. In particular,
when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (and hence private)~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and 
traditional distributed optimization~\cite{li2019convergence,smith2017federated}:
\begin{enumerate}
\item\textbf{Statistical heterogeneity:} data distributions in different devices are different.
\item\textbf{System heterogeneity:} only a subset of devices may access the central server at each time. 
\end{enumerate}

% Gaps in the theoretical understanding. 
To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success~\cite{47586}. 
Its success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.

 Existing convergence guarantees of FedAvg were either largely established on simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or provides a suboptimal 
convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} under the realistic setting is yet to be analyzed.
